Comparison of particle trajectories and collision operators for collisional transport in nonaxisymmetric plasmas

### 2014-02-05 doc/sfincsPaper/sfincsPaper.tex
Added Simakov-Helander benchmark as a new figure 6 in sfincs paper

\changed{
In the short-mean-free-path limit $\nu' \gg 1$, the ion transport and flow can be computed
analytically in terms of the parallel current \cite{AndreiPer2009}.
The transport matrix elements associated with the Fokker-Planck collision operator
may therefore be extracted from Ref. \cite{AndreiPer2009} and are summarized in Appendix \ref{apdx:SH}.
Plotted in figures \ref{fig:collisionComparison_LHD} and \ref{fig:collisionComparison_W7X} (red dashed lines),
these analytic high-collisionality limits agree quite well with the Fokker-Planck SFINCS computations
in the appropriate $\nu' \gg 1$ limit.
}

\changed{
At any collisionality, when the pure pitch-angle scattering collision operator 
and DKES trajectories are chosen, the kinetic equation
solved in SFINCS becomes identical to the one solved in the DKES code \cite{DKES1, DKES2}.
In this case, it was verified that the two codes agreed for all elements of the transport
matrix, as demonstrated in figure \ref{fig:collisionComparison_W7X}.
For this figure, the monoenergetic transport coefficients
computed by DKES have been integrated over velocity with the appropriate
weights and normalized in the same way as (\ref{eq:matrix}).
}

\begin{figure}[h!]
% Figure generated using m20140127_01_plotSFINCSFortranAndMatlabNuScansForPaper.m
% *********************************
% Previous version of figure generated using m20131208_01_plotSFINCSFortranAndMatlabNuScansForPaper.m
% High collisionality runs done in matlab, with data in
% m20130226_04_SFINCS_2013-02-26_21-21_good
% Low collisionality runs done in fortran on edison.
% *********************************
% Previous version of figure:
% Figure generated using m20131130_01_plotSFINCSCollisionOperatorComparisonForPaper.m
% Data files:
% m20131117_04_sfincs_w7x_2013-11-26_02-27_nuPrimeScan_nuPrime0_01to100
% m20130226_04_SFINCS_2013-02-26_21-21_good
% W7X run was on edison in /global/scratch2/sd/landrema/20131123-01-sfincs/20131125-02-nuPrimeScan0_01to100
% *********************************
\includegraphics{m20140127_01_plotSFINCSFortranAndMatlabNuScansForPaper_LHD.eps}
%\includegraphics{m20131208_01_plotSFINCSFortranAndMatlabNuScansForPaper_LHD.eps}
%\includegraphics{m20131130_01_plotSFINCSCollisionOperatorComparisonForPaper_LHD.eps}
\caption{(Color online) 
The ion transport matrix elements (defined in (\ref{eq:matrix})) are plotted as functions
of the collisionality (\ref{eq:nuPrime}) for LHD geometry at $E_r=0$.
SFINCS computations for three different collision operators are compared.
\changed{
Dashed red lines indicate the analytic high-collisionality limits
for Fokker-Planck collisions, discussed in appendix \ref{apdx:SH} and in Ref. \cite{AndreiPer2009},
which agree quite well with the SFINCS Fokker-Planck computations at high collisionality.
}
\label{fig:collisionComparison_LHD}}
\end{figure}

\begin{figure}[h!]
% Figure generated using m20140127_01_plotSFINCSFortranAndMatlabNuScansForPaper.m
% *********************************
% Previous version of figure generated using m20131208_01_plotSFINCSFortranAndMatlabNuScansForPaper.m
% High collisionality runs done in matlab, with data in
% m20131117_04_sfincs_w7x_2013-11-26_02-27_nuPrimeScan_nuPrime0_01to100
% Low collisionality runs done in fortran on edison.
% *********************************
% Previous version of figure:
% Figure generated using m20131130_01_plotSFINCSCollisionOperatorComparisonForPaper.m
% Data files:
% m20131117_04_sfincs_w7x_2013-11-26_02-27_nuPrimeScan_nuPrime0_01to100
% m20130226_04_SFINCS_2013-02-26_21-21_good
% W7X run was on edison in /global/scratch2/sd/landrema/20131123-01-sfincs/20131125-02-nuPrimeScan0_01to100
% *********************************
\includegraphics{m20140127_01_plotSFINCSFortranAndMatlabNuScansForPaper_W7X.eps}
%\includegraphics{m20131208_01_plotSFINCSFortranAndMatlabNuScansForPaper_W7X.eps}
%\includegraphics{m20131130_01_plotSFINCSCollisionOperatorComparisonForPaper_W7X.eps}
\caption{(Color online)
The ion transport matrix elements (defined in (\ref{eq:matrix})) are plotted as functions
of the collisionality (\ref{eq:nuPrime}) for LHD geometry at $E_r=0$.
SFINCS computations for three different collision operators are compared.
\changed{
Dashed red lines indicate the analytic high-collisionality limits
for Fokker-Planck collisions, discussed in appendix \ref{apdx:SH} and in Ref. \cite{AndreiPer2009},
which agree quite well with the SFINCS Fokker-Planck computations at high collisionality.
Also shown (black crosses) are the transport matrix elements
computed using the DKES code \cite{DKES1,DKES2}, which uses a pitch-angle scattering collision operator,
demonstrating excellent agreement with SFINCS when the latter is run with the same collision model. 
}
\label{fig:collisionComparison_W7X}}
\end{figure}

>>>>

\changed{
When the pure pitch-angle scattering collision operator 
and DKES trajectories are chosen, the kinetic equation
solved in SFINCS becomes identical to the one solved in the DKES code \cite{DKES1, DKES2}.
In this case, it was verified that the two codes agreed for all elements of the transport
matrix, as demonstrated in figure \ref{fig:collisionComparison_W7X}.
For this figure, the monoenergetic transport coefficients
computed by DKES have been integrated over velocity with the appropriate
weights and normalized in the same way as (\ref{eq:matrix}).
}

\changed{
In the short-mean-free-path limit $\nu' \gg 1$, the ion transport and flow can be computed
analytically in terms of the parallel current \cite{AndreiPer2009}.
The transport matrix elements associated with the Fokker-Planck collision operator
may therefore be extracted from Ref. \cite{AndreiPer2009} and are summarized in Appendix \ref{apdx:SH}.
Plotted in figure \ref{fig:SimakovHelanderComparison} (dashed and dot-dashed lines),
these analytic high-collisionality limits agree quite well with the Fokker-Planck SFINCS computations
in the appropriate $\nu' \gg 1$ limit.
}

\begin{figure}[h!]
% Figure generated using m20140205_01_plotSFINCSFortranAndMatlabNuScansForPaper.m
% *********************************
% Previous version of figure generated using m20131208_01_plotSFINCSFortranAndMatlabNuScansForPaper.m
% High collisionality runs done in matlab, with data in
% m20130226_04_SFINCS_2013-02-26_21-21_good
% Low collisionality runs done in fortran on edison.
% *********************************
% Previous version of figure:
% Figure generated using m20131130_01_plotSFINCSCollisionOperatorComparisonForPaper.m
% Data files:
% m20131117_04_sfincs_w7x_2013-11-26_02-27_nuPrimeScan_nuPrime0_01to100
% m20130226_04_SFINCS_2013-02-26_21-21_good
% W7X run was on edison in /global/scratch2/sd/landrema/20131123-01-sfincs/20131125-02-nuPrimeScan0_01to100
% *********************************
\includegraphics{m20140205_01_plotSFINCSFortranAndMatlabNuScansForPaper_LHD.eps}
%\includegraphics{m20131208_01_plotSFINCSFortranAndMatlabNuScansForPaper_LHD.eps}
%\includegraphics{m20131130_01_plotSFINCSCollisionOperatorComparisonForPaper_LHD.eps}
\caption{(Color online) 
The ion transport matrix elements (defined in (\ref{eq:matrix})) are plotted as functions
of the collisionality (\ref{eq:nuPrime}) for LHD geometry at $E_r=0$.
SFINCS computations for three different collision operators are compared.
\label{fig:collisionComparison_LHD}}
\end{figure}

\begin{figure}[h!]
% Figure generated using m20140205_01_plotSFINCSFortranAndMatlabNuScansForPaper.m
% *********************************
% Previous version of figure generated using m20131208_01_plotSFINCSFortranAndMatlabNuScansForPaper.m
% High collisionality runs done in matlab, with data in
% m20131117_04_sfincs_w7x_2013-11-26_02-27_nuPrimeScan_nuPrime0_01to100
% Low collisionality runs done in fortran on edison.
% *********************************
% Previous version of figure:
% Figure generated using m20131130_01_plotSFINCSCollisionOperatorComparisonForPaper.m
% Data files:
% m20131117_04_sfincs_w7x_2013-11-26_02-27_nuPrimeScan_nuPrime0_01to100
% m20130226_04_SFINCS_2013-02-26_21-21_good
% W7X run was on edison in /global/scratch2/sd/landrema/20131123-01-sfincs/20131125-02-nuPrimeScan0_01to100
% *********************************
\includegraphics{m20140205_01_plotSFINCSFortranAndMatlabNuScansForPaper_W7X.eps}
%\includegraphics{m20131208_01_plotSFINCSFortranAndMatlabNuScansForPaper_W7X.eps}
%\includegraphics{m20131130_01_plotSFINCSCollisionOperatorComparisonForPaper_W7X.eps}
\caption{(Color online)
The ion transport matrix elements (defined in (\ref{eq:matrix})) are plotted as functions
of the collisionality (\ref{eq:nuPrime}) for W7-X geometry at $E_r=0$.
SFINCS computations for three different collision operators are compared.
\changed{
Also shown (black crosses) are the transport matrix elements
computed using the DKES code \cite{DKES1,DKES2}, which uses a pitch-angle scattering collision operator,
demonstrating excellent agreement with SFINCS when the latter is run with the same collision model. 
}
\label{fig:collisionComparison_W7X}}
\end{figure}

\begin{figure}[h!]
% Figure generated using m20140205_02_plotSFINCSVsSimakovHelanderLimitForPaper.m
\includegraphics{m20140205_02_plotSFINCSVsSimakovHelanderLimitForPaper.eps}
\caption{(Color online)
\changed{
The ion transport matrix elements (defined in (\ref{eq:matrix})) are plotted as functions
of the collisionality (\ref{eq:nuPrime}) for LHD and W7-X geometry at $E_r=0$.
SFINCS results shown were computed using the linearized Fokker-Planck collision operator
(so the solid curves here are
identical to the red curves in figures \ref{fig:collisionComparison_LHD}-\ref{fig:collisionComparison_W7X}.)
Dashed and dot-dashed lines indicate the analytic high-collisionality limits
for Fokker-Planck collisions, discussed in Ref. \cite{AndreiPer2009} and in appendix \ref{apdx:SH},
which agree quite well with the SFINCS computations at high collisionality.
}
\label{fig:SimakovHelanderComparison}}
\end{figure}
